---
title: "Class10 - CANDY"
author: "Madeleine Duquette"
format: pdf
date: 10/28/22
---

# Background

In this mini-project we will examine 538 Halloween Candy data. What is your favorite candy? What is nougat anyway? 

First step is to read the data...
```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```
# Q1. How many different candy types are in this dataset?
```{r}
nrow(candy)
```

# Q2. How many fruity candy types are in the dataset?
```{r}
sum(candy$fruity)
```
# Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
```{r}
#rownames(candy)
#Sour Patch kids are the best...
candy["Sour Patch Kids", ]$winpercent
```

# Q4. What is the winpercent value for “Kit Kat”?
```{r}
candy["Kit Kat", ]$winpercent
```

# Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```


```{r}
skimr::skim(candy)
```
# Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
winpercent is on a different scale

# Q7. What do you think a zero and one represent for the candy$chocolate column?
Whether or not the candy is a chocolate

# Q8. Plot a histogram of winpercent values
```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)
ggplot(candy, aes(x=winpercent)) + geom_histogram(bins=10, col="red", fill="turquoise")
```

# Q9. Is the distribution of winpercent values symmetrical?
No it does not appear to be symmetrical

# Q10. Is the center of the distribution above or below 50%?
Below

# Q11. On average is chocolate candy higher or lower ranked than fruit candy?
```{r}
chocolate <- candy$chocolate == 1
choco_score <- mean(candy[chocolate,]$winpercent)

fruity <- candy$fruity == 1
fruity_score <- mean(candy[fruity,]$winpercent)

choco_score > fruity_score
```

# Q12. Is this difference statistically significant?
```{r}
t.test(candy[chocolate,]$winpercent, candy[fruity,]$winpercent)
```
Yes!


# The base R `sort()` and `order()` functions are very useful!

# Q13. What are the five least liked candy types in this set?
```{r}
library(dplyr)
candy %>% 
  arrange(winpercent) %>% 
  head(5)
```

# Q14. What are the top 5 all time favorite candy types out of this set?
```{r}
library(dplyr)
candy %>% 
  arrange(desc(winpercent)) %>% 
  head(5)
```

# Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

# Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?
```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

```{r}
ggsave("mybarplot.png")
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$chocolate)] = "chocolate"
# my_cols[as.logical(candy$bar & candy$chocolate)] = "blue"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols)
```

# Q17. What is the worst ranked chocolate candy?
Sixlets

# Q18. What is the best ranked fruity candy?
Starburst

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=2, max.overlaps = 20)
```
# Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
Reese's Mini

# Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
```{r}
candy %>% 
  arrange(desc(pricepercent)) %>% 
  head(5)
```
Nik L Nip is the least popular

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
# Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
Chocolate and fruity

# Q23. Similarly, what two variables are most positively correlated?
Chocolate and winpercent

## PCA

The main function that is always there for us is `prcomp()`. It has an important argument that is set to `scale=FALSE` but needs to be set to `TRUE` in this case because our data is not on the same scale.
```{r}
pca <- prcomp(candy, scale=T)
summary(pca)
```
My PCA plot (aka) PC1 vs PC2 score plot.
```{r}
plot(pca$x[,1], pca$x[,2], col=my_cols, pch=16)
```
I will make a "nicer" plot with ggplot
ggplot only works with data.frames as input so I need to make one...

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
library(ggrepel)
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=2, col=my_cols, max.overlaps = 25)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

# Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
Fruity, hard, and pluribus. It makes sense these are all grouping together in one direction because that's the typical form for fruity candies... Fruit candies thrive on variety! And these characteristics are not common for chocolates.


